---
title: "Data Analyst Portfolio"
subtitle: "American University SIS 750 | Spring 2026"
author: "Brendan Joyce"
date: "2026-05-06"
format:
  revealjs:
    theme: night
    slide-number: h.v
    transition: fade
    transition-speed: fast
    background-transition: zoom
    embed-resources: true
    center: true
    margin: 0.05
    width: 1400
    height: 800
    auto-animate: true
    auto-animate-easing: ease-in-out
    auto-animate-duration: 600
    controls: true
    progress: true
    hash: true
    keyboard: true
    highlight-theme: dracula
    pdf-separate-fragments: true
    plugins:
      - search
      - zoom
      - notes
      - mathjax
execute:
  echo: false
  warning: false
  message: false
  fig.align: center
---

```{r}

library(tidyverse)   
library(knitr)       
library(scales) 
library(ggtext)
library(ggplot2)
library(maps)
library(devtools)
library(tidyr)
library(dplyr)
library(lubridate)
library(scales)
library(patchwork)
library(stringr)
library(viridisLite)


```

## PS2: Who Responded to the Invasion? {.center .smaller}

Great powers and regional actors drove the response — reflecting renewed geopolitical competition

```{r}
c_intl_response_to_ukraine_crisis <- read_csv("intl_response_to_ukraine_war.csv")


nato_dates_split <- c_intl_response_to_ukraine_crisis %>%
  mutate(
    date = as.Date(mdy_hms(`Start Date`)),
    nato_status = if_else(
      `Event Country` %in% c(
        "Belgium","Bulgaria","Canada","Czech Republic","Estonia","France","Germany",
        "Greece","Italy","Lithuania","Netherlands","Norway","Poland","Portugal",
        "Romania","Slovakia","Spain","Turkey","United Kingdom","United States"
      ),
      "NATO", "Non-NATO"
    ),
    cutoff = case_when(
      date >= as.Date("2022-02-24") ~ "post invasion",
      date <  as.Date("2022-02-24") ~ "pre invasion"
    )
  )

pre_invasion_counts <- nato_dates_split %>%
  filter(cutoff == "pre invasion") %>%
  count(`Event Country`, name = "pre_invasion_total_responses")

post_invasion_counts <- nato_dates_split %>%
  filter(cutoff == "post invasion") %>%
  count(`Event Country`, name = "post_invasion_total_responses")

country_response_counts <- pre_invasion_counts %>%
  full_join(post_invasion_counts, by = "Event Country") %>%
  replace_na(list(
    pre_invasion_total_responses = 0,
    post_invasion_total_responses = 0
  ))

country_proportions_pct <- country_response_counts %>%
  mutate(
    total_pre  = sum(pre_invasion_total_responses),
    total_post = sum(post_invasion_total_responses),
    pre_pct   = 100 * pre_invasion_total_responses / total_pre,
    post_pct  = 100 * post_invasion_total_responses / total_post,
    pct_change = post_pct - pre_pct
  ) %>%
  select(`Event Country`, pre_pct, post_pct, pct_change)

descending_pcts_post <- country_proportions_pct %>% 
   arrange(desc(post_pct)) %>% 
  select(`Event Country`, post_pct, pct_change)

descending_pcts_pre <-  country_proportions_pct %>% 
  arrange(desc(pre_pct)) %>% 
  select(`Event Country`, pre_pct, pct_change)


responses_before_after_invasion_plot <- country_response_counts %>%
  pivot_longer(
     cols = c(pre_invasion_total_responses, post_invasion_total_responses),
    names_to = "period",
    values_to = "count"
  )

# Nato versus Non-Nato
nato_countries <- c("Belgium","Bulgaria","Canada","Czech Republic","Estonia","France","Germany",
                    "Greece","Italy","Lithuania","Netherlands","Norway","Poland","Portugal",
                    "Romania","Slovakia","Spain","Turkey","United Kingdom","United States")


top_nato_countries <- c("United States", "United Kingdom", "Germany")


top_non_nato_countries <- c("Ukraine", "Russian Federation", "India")


top_countries_list <- c(top_nato_countries, top_non_nato_countries)

#  Total responses by week 
all_timeline_data <- c_intl_response_to_ukraine_crisis %>%
  mutate(
    date = as.Date(mdy_hms(`Start Date`)),
    week = floor_date(date, "week"),
    nato_status = if_else(`Event Country` %in% nato_countries, "NATO", "Non-NATO")
  ) %>%
  count(`Event Country`, week, nato_status) %>%
  arrange(`Event Country`, week) %>%
  group_by(`Event Country`) %>%
  mutate(cumulative = cumsum(n)) %>%
  ungroup()

# Average cumulative responses across  NATO 
nato_avg_timeline <- all_timeline_data %>%
  filter(nato_status == "NATO") %>%
  group_by(week) %>%
  summarise(avg_cumulative = mean(cumulative), .groups = "drop")

# Average cumulative responses across Non-NATO countries
non_nato_avg_timeline <- all_timeline_data %>%
  filter(nato_status == "Non-NATO") %>%
  group_by(week) %>%
  summarise(avg_cumulative = mean(cumulative), .groups = "drop")

# Total responses by week for top 3 countries 
timeline_data <- c_intl_response_to_ukraine_crisis %>%
  filter(`Event Country` %in% top_countries_list) %>%
  mutate(
    date = as.Date(mdy_hms(`Start Date`)),
    week = floor_date(date, "week")
  ) %>%
  count(`Event Country`, week) %>%
   arrange(`Event Country`, week) %>%
  group_by(`Event Country`) %>%
    mutate(cumulative = cumsum(n)) %>%
  ungroup() %>%
  mutate(nato_status = if_else(`Event Country` %in% nato_countries, "NATO", "Non-NATO"))

date_range <- c_intl_response_to_ukraine_crisis %>%
  mutate(date = as.Date(mdy_hms(`Start Date`))) %>%
  summarise(min_date = min(date, na.rm = TRUE), max_date = max(date, na.rm = TRUE))

y_max <- max(timeline_data$cumulative)


timeline_nato <- timeline_data %>% filter(nato_status == "NATO")

timeline_non_nato <- timeline_data %>% filter(nato_status == "Non-NATO")

#  NATO plot 
nato_plot <- ggplot(timeline_nato, aes(x = week, y = cumulative, color = `Event Country`, shape = `Event Country`)) +
  geom_line(data = nato_avg_timeline, aes(x = week, y = avg_cumulative), 
            color = "#757575", linewidth = 1.2, linetype = "dotted", alpha = 0.6, inherit.aes = FALSE) +
  annotate("text", x = max(nato_avg_timeline$week), y = max(nato_avg_timeline$avg_cumulative) * 1.05,
           label = "NATO Avg", color = "#757575", size = 2.5, hjust = 1, fontface = "italic", family = "sans") +
  geom_vline(xintercept = as.Date("2022-02-24"), linetype = "dotted", color = "#d32f2f", linewidth = 0.8, alpha = 0.7) +
  annotate("text", x = as.Date("2022-02-24"), y = y_max * 0.95, 
           label = "INVASION", color = "#d32f2f", fontface = "bold", size = 3, hjust = -0.1, family = "sans") +
  geom_line(linewidth = 1.3, alpha = 0.95) +
  geom_point(size = 2.2, alpha = 0.9, stroke = 0.5) +
  scale_color_manual(
    values = c(
      "United States" = "#1565c0",
      "United Kingdom" = "#e65100", 
      "Germany" = "#6a1b9a"
    )
  ) +
  scale_x_date(
    limits = c(date_range$min_date - 2, date_range$max_date + 2),
    date_breaks = "2 weeks",
    date_labels = "%d %b",
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_continuous(
    limits = c(0, y_max),
    expand = expansion(mult = c(0, 0.05)),
    breaks = seq(0, 200, by = 20)
  ) +
  labs(
    title = "NATO MEMBER RESPONSES",
    x = NULL,
    y = "Cumulative Responses"
  ) +
  theme_minimal(base_size = 11, base_family = "sans") +
  theme(
    plot.title = element_text(face = "bold", size = 11, hjust = 0.5, margin = margin(b = 8), 
                              color = "#212121", family = "sans"),
    axis.text.x = element_text(size = 8.5, color = "#424242", angle = 0, hjust = 0.5, family = "sans"),
    axis.text.y = element_text(size = 8.5, color = "#424242", family = "sans"),
    axis.title.y = element_text(size = 9.5, face = "bold", color = "#212121", family = "sans"),
    legend.position = "bottom",
    legend.text = element_text(size = 8.5, family = "sans"),
    legend.title = element_blank(),
     legend.key.height = unit(0.4, "cm"),
     legend.key.width = unit(0.8, "cm"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "#e0e0e0", linewidth = 0.25),
    panel.border = element_rect(color = "#bdbdbd", fill = NA, linewidth = 0.5),
    plot.background = element_rect(fill = "#fafafa", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(10, 5, 10, 10)
  ) +
  guides(color = guide_legend(nrow = 1, override.aes = list(size = 3)))

# Non-NATO visual
non_nato_plot <- ggplot(timeline_non_nato, aes(x = week, y = cumulative, color = `Event Country`, shape = `Event Country`)) +
  geom_line(data = non_nato_avg_timeline, aes(x = week, y = avg_cumulative), 
            color = "#757575", linewidth = 1.2, linetype = "dotted", alpha = 0.6, inherit.aes = FALSE) +
  annotate("text", x = max(non_nato_avg_timeline$week), y = max(non_nato_avg_timeline$avg_cumulative) * 1.05,
           label = "Non-NATO Avg", color = "#757575", size = 2.5, hjust = 1, fontface = "italic", family = "sans") +
  geom_vline(xintercept = as.Date("2022-02-24"), linetype = "dotted", color = "#d32f2f", linewidth = 0.8, alpha = 0.7) +
  annotate("text", x = as.Date("2022-02-24"), y = y_max * 0.95, 
           label = "INVASION", color = "#d32f2f", fontface = "bold", size = 3, hjust = -0.1, family = "sans") +
  geom_line(linewidth = 1.3, alpha = 0.95) +
  geom_point(size = 2.2, alpha = 0.9, stroke = 0.5) +
  scale_color_manual(
    values = c(
      "Ukraine" = "#c62828",
      "Russian Federation" = "#283593",
      "India" = "#00695c"
    )
  ) +
  scale_x_date(
    limits = c(date_range$min_date - 2, date_range$max_date + 2),
    date_breaks = "2 weeks",
    date_labels = "%d %b",
    expand = expansion(mult = c(0.02, 0.02))
  ) +
  scale_y_continuous(
    limits = c(0, y_max),
    expand = expansion(mult = c(0, 0.05)),
    breaks = seq(0, 200, by = 20)
  ) +
  labs(
    title = "NON-NATO COUNTRY RESPONSES",
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 11, base_family = "sans") +
  theme(
    plot.title = element_text(face = "bold", size = 11, hjust = 0.5, margin = margin(b = 8), 
                              color = "#212121", family = "sans"),
    axis.text.x = element_text(size = 8.5, color = "#424242", angle = 0, hjust = 0.5, family = "sans"),
    axis.text.y = element_text(size = 8.5, color = "#424242", family = "sans"),
    legend.position = "bottom",
    legend.text = element_text(size = 8.5, family = "sans"),
    legend.title = element_blank(),
    legend.key.height = unit(0.4, "cm"),
    legend.key.width = unit(0.8, "cm"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "#e0e0e0", linewidth = 0.25),
    panel.border = element_rect(color = "#bdbdbd", fill = NA, linewidth = 0.5),
    plot.background = element_rect(fill = "#fafafa", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(10, 10, 10, 5)
  ) +
  guides(color = guide_legend(nrow = 1, override.aes = list(size = 3)))

pre_post_invasion_plot <- nato_plot | non_nato_plot

# Combined plots 
pre_post_invasion_plot <- pre_post_invasion_plot +
  plot_annotation(
    title = NULL,
    caption = "Source: Janes Intelligence Review (Feb-Apr 2022). Countries selected to show major responders across alliance structures.",
    theme = theme(
      plot.caption = element_text(size = 8.5, color = "#424242", hjust = 0.5, margin = margin(t = 5), family = "sans", lineheight = 1.3)
    )
  )


nato <- c_intl_response_to_ukraine_crisis %>%
  mutate(
    date = as.Date(mdy_hms(`Start Date`)),
    nato_status = if_else(
      `Event Country` %in% c(
        "Belgium","Bulgaria","Canada","Czech Republic","Estonia","France","Germany",
        "Greece","Italy","Lithuania","Netherlands","Norway","Poland","Portugal",
        "Romania","Slovakia","Spain","Turkey","United Kingdom","United States"
      ),
      "NATO", "Non-NATO")
    )

events_by_day <- nato %>%
  group_by(date, `Event Type V2`) %>%
  summarise(count = n(), .groups = "drop")

events_by_day_graph <- 
  ggplot(
    events_by_day, aes(x = date, y = count,
                       color = `Event Type V2`,
                       fill  = `Event Type V2`)
  ) +
  geom_area(alpha = 0.2, position = "identity") +
  geom_line(linewidth = 0.9) +
  labs(
    x     = NULL,
    y     = "Number of Events",
    color = "Action Type",
    fill  = "Action Type"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position  = "bottom",
    legend.text      = element_text(size = 8),
    panel.grid.minor = element_blank(),
    axis.text.x      = element_text(angle = 30, hjust = 1, size = 7)
  )


border_theme <- theme(
  plot.background = element_rect(fill = "white", colour = "black", linewidth = 0.8)
)


world_responses <- pre_post_invasion_plot

world_responses

```

## Global Response Intensity {.center .smaller}

Map Indicates that responses to Ukraine were mostly between Great Powers
```{r}
#| label: ps2-response-intensity-map

library(maps)
library(viridisLite)

# Response counts by country and type
response_by_country_type <- c_intl_response_to_ukraine_crisis %>%
  count(`Event Country`, `Event Type V2`) %>%
  pivot_wider(names_from = `Event Type V2`, values_from = n, values_fill = 0) %>%
  mutate(
    total_responses = Statement + Operation + Action + Judicial + Attack,
    primary_intensity = Statement + Action,
    secondary_intensity = Operation + Attack + Judicial,
    response_strategy = case_when(
      (Operation + Attack) > (Statement + Action) ~ "Kinetic",
      Statement > Action ~ "Diplomatic",
      TRUE ~ "Mixed"
    )
  )

world_map <- map_data("world")

# I covered all countries in case future analysts want to delve into other regions of interest
country_mapping <- tribble(
  ~`Event Country`, ~region,
  "United States", "USA",
  "United Kingdom", "UK",
   "Russian Federation", "Russia",
  "Germany", "Germany",
  "France", "France",
  "Italy", "Italy",
  "Spain", "Spain",
  "Poland", "Poland",
  "Ukraine", "Ukraine",
  "Turkey", "Turkey",
  "Canada", "Canada",
  "Netherlands", "Netherlands",
  "Belgium", "Belgium",
  "Greece", "Greece",
  "Portugal", "Portugal",
  "Romania", "Romania",
  "Bulgaria", "Bulgaria",
  "Czech Republic", "Czech Republic",
  "Hungary", "Hungary",
  "Slovakia", "Slovakia",
  "Lithuania", "Lithuania",
  "Latvia", "Latvia",
  "Estonia", "Estonia",
  "Norway", "Norway",
  "Sweden", "Sweden",
  "Finland", "Finland",
  "Denmark", "Denmark",
  "Austria", "Austria",
  "Switzerland", "Switzerland",
  "Ireland", "Ireland",
  "Croatia", "Croatia",
  "Slovenia", "Slovenia",
  "Serbia", "Serbia",
  "Albania", "Albania",
  "North Macedonia", "North Macedonia",
  "Bosnia and Herzegovina", "Bosnia and Herzegovina",
  "Montenegro", "Montenegro",
  "Moldova", "Moldova",
  "Georgia", "Georgia",
  "Armenia", "Armenia",
  "Azerbaijan", "Azerbaijan",
  "Japan", "Japan",
  "South Korea", "South Korea",
  "Australia", "Australia",
  "New Zealand", "New Zealand",
  "China", "China",
  "India", "India",
  "Singapore", "Singapore",
  "Taiwan", "Taiwan",
  "Israel", "Israel",
  "Saudi Arabia", "Saudi Arabia",
  "United Arab Emirates", "UAE",
  "Qatar", "Qatar",
   "Iran", "Iran",
  "Egypt", "Egypt",
  "South Africa", "South Africa",
  "Brazil", "Brazil",
  "Argentina", "Argentina",
  "Chile", "Chile",
  "Mexico", "Mexico",
  "Colombia", "Colombia"
)


response_map_data <- response_by_country_type %>%
  left_join(country_mapping, by = "Event Country") %>%
  filter(!is.na(region))

world_response <- world_map %>%
  left_join(response_map_data, by = "region")

# gradient map (legend is logged)
ggplot(world_response, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = total_responses), color = "grey30", linewidth = 0.1) +
  geom_polygon(data = world_response %>% filter(response_strategy == "Kinetic"),
               aes(x = long, y = lat, group = group),
               fill = NA, color = "#ef4444", linewidth = 1.2) +
  scale_fill_gradientn(
    colors = c("#ffffff", "#fee5d9", "#fcbba1", "#fc9272", "#fb6a4a", "#ef3b2c", "#cb181d", "#99000d"),
    na.value = "grey20",
    name = "Total Responses",
    breaks = c(1, 10, 25, 50, 100, 150),
    trans = "log1p",
    labels = c("1", "10", "25", "50", "100", "150+")
  ) +
  coord_fixed(1.3, ylim = c(-55, 75), expand = FALSE) +
  labs(
    title = "Other Non-NATO Responses Indicate Clear Global Concern-India, Argentina, Japan",
    caption = "Source: Janes Intelligence Review (Feb-Dec 2022). Note: Russia, Ukraine, Georgia, UK employed Operations/Attacks, but less so than actions & statements"
  ) +
  theme_void(base_size = 12) +
  theme(
    plot.background = element_rect(fill = "grey10", color = NA),
    panel.background = element_rect(fill = "grey10", color = NA),
     plot.title = element_text(color = "white", face = "bold", size = 13, hjust = 0.5, margin = margin(b = 5, t = 5)),
    plot.subtitle = element_text(color = "grey80", size = 10, hjust = 0.5, margin = margin(b = 10)),
    plot.caption = element_text(color = "grey70", size = 8, hjust = 0.5, margin = margin(t = 8, b = 5)),
    legend.position = "bottom",
    legend.title = element_text(color = "white", size = 10, face = "bold"),
    legend.text = element_text(color = "white", size = 9),
    legend.background = element_rect(fill = "grey10", color = NA),
     legend.key.width = unit(2.5, "cm"),
     legend.key.height = unit(0.4, "cm"),
     plot.margin = margin(5, 5, 5, 5)
  )
```

---

## Response Type by Country {.center .smaller}

States employed different response strategies 

```{r}
#| label: ps2-stacked-bar-chart
#| fig-width: 12
#| fig-height: 6

# Top responding countries and types of responses
top_countries <- country_response_counts %>%
  mutate(total = pre_invasion_total_responses + post_invasion_total_responses) %>%
  slice_max(order_by = total, n = 12) %>%
  pull(`Event Country`)

event_type_breakdown <- c_intl_response_to_ukraine_crisis %>%
  filter(`Event Country` %in% top_countries) %>%
  count(`Event Country`, `Event Type V2`) %>%
  group_by(`Event Country`) %>%
  mutate(
    total = sum(n),
    proportion = n / total,
    pct_label = if_else(proportion >= 0.08, paste0(round(proportion * 100), "%"), "")
  ) %>%
  ungroup() %>%
  mutate(
    `Event Country` = fct_reorder(`Event Country`, total),
    `Event Type V2` = factor(`Event Type V2`, levels = c("Statement", "Operation", "Judicial", "Attack", "Action"))
  )

# Horizontal Bar Chart
ggplot(event_type_breakdown, aes(x = `Event Country`, y = n, fill = `Event Type V2`)) +
  annotate("rect", xmin = 0.5, xmax = which(levels(event_type_breakdown$`Event Country`) == "United States") + 0.5,
           ymin = 0, ymax = Inf, fill = "#e3f2fd", alpha = 0.2) +
  geom_col(width = 0.8, color = "white", linewidth = 0.4) +
  geom_text(aes(label = pct_label), 
            position = position_stack(vjust = 0.5),
            size = 3, fontface = "bold", color = "white") +
  scale_fill_manual(
    values = c(
      "Statement" = "#2563eb",    
      "Operation" = "#ea580c",    
      "Action" = "#059669",       
      "Judicial" = "#7c3aed",     
      "Attack" = "#dc2626"        
    ),
    name = "Response Type"
  ) +
  scale_y_continuous(
    expand = expansion(mult = c(0, 0.05)),
    breaks = seq(0, 200, by = 25)
  ) +
  labs(
    title = "Most States Undertook a Restrained & Diplomatic Approach in Response",
    x = NULL,
    y = "Total Number of Recorded Responses",
    caption = "Source: Janes Intelligence Review (Feb 2022 - Dec 2022). Top 12 responding countries shown. Percentages displayed for response types ≥8%."
  ) +
  coord_flip() +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0, margin = margin(b = 5)),
    plot.subtitle = element_text(size = 11, hjust = 0, color = "grey30", margin = margin(b = 15)),
    legend.position = "bottom",
    legend.text = element_text(size = 10),
    legend.title = element_text(size = 11, face = "bold"),
    legend.box.background = element_rect(color = "grey80", linewidth = 0.5),
    legend.margin = margin(t = 10),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(color = "grey85", linewidth = 0.4),
    plot.caption = element_text(size = 8, color = "black", hjust = 0, margin = margin(t = 10)),
    axis.text.y = element_text(size = 11, face = "bold", color = "grey20"),
     axis.text.x = element_text(size = 10, color = "grey30"),
     axis.title.x = element_text(size = 11, face = "bold", margin = margin(t = 10)),
     plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(10, 10, 10, 10)
  )

```

---

## PS4: Diffusion of Armed UAV Adoption {.center .smaller}

Early adopters in the U.S., Israel, U.K., Iran, and China 

```{r}
#| label: ps4-uav-map

library(tidyverse)
library(maps)
library(viridisLite)

cow_long <-  read_csv('cow_long.csv')

uav_adoption_all <- cow_long |>
  filter(techtype == "Armed UAVs", !is.na(use), use %in% c(1, 9)) |>
  group_by(statename) |>
  summarise(adoption_year = min(year), .groups = "drop")

world_map <- map_data("world")

country_mapping <- tribble(
  ~statename, ~region,
  "United States of America", "USA",
  "United Kingdom", "UK",
  "Russia", "Russia",
  "China", "China",
  "Turkey", "Turkey",
  "Iran", "Iran",
  "Israel", "Israel",
  "Pakistan", "Pakistan",
  "United Arab Emirates", "UAE",
  "Saudi Arabia", "Saudi Arabia",
  "Nigeria", "Nigeria",
  "France", "France",
  "Germany", "Germany",
  "Italy", "Italy",
  "Spain", "Spain",
  "India", "India",
  "South Korea", "South Korea",
  "Japan", "Japan",
  "Australia", "Australia",
  "Canada", "Canada",
  "Brazil", "Brazil",
  "South Africa", "South Africa",
  "Egypt", "Egypt",
  "Morocco", "Morocco",
  "Algeria", "Algeria",
  "Tunisia", "Tunisia",
  "Jordan", "Jordan",
  "Iraq", "Iraq",
  "Azerbaijan", "Azerbaijan",
  "Kazakhstan", "Kazakhstan",
  "Uzbekistan", "Uzbekistan",
  "Indonesia", "Indonesia",
  "Malaysia", "Malaysia",
  "Singapore", "Singapore",
  "Thailand", "Thailand",
  "Vietnam", "Vietnam",
  "Poland", "Poland",
  "Ukraine", "Ukraine",
  "Greece", "Greece",
  "Netherlands", "Netherlands",
  "Belgium", "Belgium",
  "Sweden", "Sweden",
  "Norway", "Norway",
  "Denmark", "Denmark",
  "Finland", "Finland",
  "Austria", "Austria",
  "Switzerland", "Switzerland",
  "Portugal", "Portugal",
  "Czech Republic", "Czech Republic",
  "Hungary", "Hungary",
  "Romania", "Romania",
  "Bulgaria", "Bulgaria",
  "Serbia", "Serbia",
  "Croatia", "Croatia",
  "Slovenia", "Slovenia",
  "Slovakia", "Slovakia",
  "Lithuania", "Lithuania",
  "Latvia", "Latvia",
  "Estonia", "Estonia",
  "Ethiopia", "Ethiopia",
  "Kenya", "Kenya",
  "Libya", "Libya",
  "Myanmar", "Myanmar",
  "Philippines", "Philippines",
  "Taiwan", "Taiwan",
  "Venezuela", "Venezuela",
  "Argentina", "Argentina",
  "Chile", "Chile",
  "Colombia", "Colombia",
  "Peru", "Peru",
  "Mexico", "Mexico",
  "South Sudan", "South Sudan",
  "Sudan", "Sudan",
   "Syria", "Syria",
   "Yemen", "Yemen",
  "Afghanistan", "Afghanistan",
  "Bangladesh", "Bangladesh",
  "Nepal", "Nepal",
  "Sri Lanka", "Sri Lanka",
  "Armenia", "Armenia",
  "Georgia", "Georgia",
  "Belarus", "Belarus",
  "Moldova", "Moldova",
  "Albania", "Albania",
   "Bosnia and Herzegovina", "Bosnia and Herzegovina",
   "North Macedonia", "North Macedonia",
   "Montenegro", "Montenegro",
  "Cyprus", "Cyprus",
  "Lebanon", "Lebanon",
  "Kuwait", "Kuwait",
  "Bahrain", "Bahrain",
  "Qatar", "Qatar",
  "Oman", "Oman",
  "Burkina Faso", "Burkina Faso",
  "Chad", "Chad",
  "Democratic Republic of the Congo", "Democratic Republic of the Congo",
  "Djibouti", "Djibouti",
   "Kyrgyzstan", "Kyrgyzstan",
  "Tajikistan", "Tajikistan",
  "Turkmenistan", "Turkmenistan",
  "Mali", "Mali",
  "Niger", "Niger",
  "Somalia", "Somalia",
  "Uganda", "Uganda",
  "Zimbabwe", "Zimbabwe"
)

uav_map_data <- uav_adoption_all |>
  left_join(country_mapping, by = "statename") |>
  filter(!is.na(region))

world_uav <- world_map |>
  left_join(uav_map_data, by = "region")

# A Gradient Map of UAV Diffusion
ggplot(world_uav, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = adoption_year), color = "grey30", linewidth = 0.1) +
  scale_fill_gradientn(
    colors = c("#2166ac", "#4393c3", "#92c5de", "#d1e5f0", "#fddbc7", "#f4a582", "#d6604d", "#b2182b"),
    na.value = "grey20",
    name = "Adoption Year",
    breaks = seq(1990, 2023, by = 5),
    labels = seq(1990, 2023, by = 5),
    limits = c(1990, 2023)
  ) +
  coord_fixed(1.3, ylim = c(-55, 83), expand = FALSE) +
  labs(
    title = "UAV Diffusion Eminates from the Middle East (Iran) and Asia (China)",
    caption = "Source: COW Arms Technology Dataset v1.0 (2025). Blue = early adopters, Red = recent adopters."
  ) +
  theme_void(base_size = 13) +
  theme(
     plot.background = element_rect(fill = "grey10", color = NA),
     panel.background = element_rect(fill = "grey10", color = NA),
    plot.title = element_text(color = "white", face = "bold", size = 13, hjust = 0.5, margin = margin(b = 10, t = 5)),
    plot.caption = element_text(color = "grey70", size = 7.5, hjust = 0.5, margin = margin(t = 8, b = 5)),
    legend.position = "bottom",
    legend.title = element_text(color = "white", size = 10),
    legend.text = element_text(color = "white", size = 9),
    legend.background = element_rect(fill = "grey10", color = NA),
    legend.key.width = unit(2, "cm"),
    plot.margin = margin(2, 2, 2, 2)
  )
```

---

## Adoption of UAVs vs. Legacy Platforms {.center .smaller}

Armed UAV adoption is spreading faster than any other advanced military technology 

```{r}
#| label: ps4-multi-tech-adoption-rates


# This code here focuses on armed UAVs, 5th Gen Fighters, 2nd Gen Combat Helicopters, 3rd Gen tanks

tech_adoption_multi <- cow_long |>
  filter(
    (techtype == "Armed UAVs") |
    (techtype == "Fighter aircraft" & techname == "Fifth gen. jet fighter") |
    (techtype == "Attack helicopters" & techname == "Second gen. helicopter") |
    (techtype == "Tanks" & techname == "Third gen. MBT"),
    !is.na(use),
    use == 1
  ) |>
  mutate(
    tech_label = case_when(
      techtype == "Armed UAVs" ~ "Armed UAVs",
      techname == "Fifth gen. jet fighter" ~ "5th Gen Fighters",
      techname == "Second gen. helicopter" ~ "2nd Gen Combat Helicopters",
      techname == "Third gen. MBT" ~ "3rd Gen Main Battle Tanks",
      TRUE ~ techname
    )
  ) |>
  group_by(tech_label, statename) |>
  summarise(adoption_year = min(year), .groups = "drop") |>
  group_by(tech_label, adoption_year) |>
  summarise(n = n(), .groups = "drop") |>
  arrange(tech_label, adoption_year) |>
  group_by(tech_label) |>
  mutate(cumulative_adopters = cumsum(n)) |>
  ungroup()


ggplot(tech_adoption_multi, aes(x = adoption_year, y = cumulative_adopters, 
                                 color = tech_label, fill = tech_label, shape = tech_label)) +
  annotate("rect", xmin = 2001, xmax = 2003, ymin = 0, ymax = Inf, 
           fill = "#fee5d9", alpha = 0.3) +
   annotate("rect", xmin = 2014, xmax = 2016, ymin = 0, ymax = Inf, 
           fill = "#fcbba1", alpha = 0.3) +
  annotate("rect", xmin = 2020, xmax = 2022, ymin = 0, ymax = Inf, 
           fill = "#fc9272", alpha = 0.3) +
  geom_area(alpha = 0.25, position = "identity", linewidth = 0) +
  geom_line(linewidth = 1.2, alpha = 0.9) +
  geom_point(data = tech_adoption_multi %>% filter(adoption_year %% 1 == 0),
             size = 2.5, alpha = 0.95, stroke = 0.8) +
   geom_vline(xintercept = 2001, linetype = "dotted", color = "#cb181d", linewidth = 0.8, alpha = 0.7) +
  geom_vline(xintercept = 2014, linetype = "dotted", color = "#cb181d", linewidth = 0.8, alpha = 0.7) +
  geom_vline(xintercept = 2020, linetype = "dotted", color = "#cb181d", linewidth = 0.8, alpha = 0.7) +
  annotate("text", x = 2001, y = 72, label = "9/11 & GWOT", 
           size = 3.5, fontface = "bold", color = "#67000d") +
  annotate("segment", x = 2001, xend = 2001, y = 70, yend = 65, 
           arrow = arrow(length = unit(0.2, "cm")), color = "#67000d", linewidth = 0.6) +
  annotate("text", x = 2014, y = 72, label = "Crimea Annexed", 
           size = 3.5, fontface = "bold", color = "#67000d") +
  annotate("segment", x = 2014, xend = 2014, y = 70, yend = 65, 
           arrow = arrow(length = unit(0.2, "cm")), color = "#67000d", linewidth = 0.6) +
  annotate("text", x = 2020, y = 70, label = "Nagorno-Karabakh\nDrone War", 
           size = 3.5, fontface = "bold", color = "#67000d", lineheight = 0.9) +
  annotate("segment", x = 2020, xend = 2020, y = 65, yend = 60, 
           arrow = arrow(length = unit(0.2, "cm")), color = "#67000d", linewidth = 0.6) +
  scale_color_manual(
    values = c(
      "Armed UAVs" = "#E69F00",              
      "5th Gen Fighters" = "#0173b2",        
      "2nd Gen Combat Helicopters" = "#029e73", 
      "3rd Gen Main Battle Tanks" = "#cc78bc"   
    ),
    name = "Military Technology"
  ) +
  scale_fill_manual(
    values = c(
      "Armed UAVs" = "#E69F00",              
      "5th Gen Fighters" = "#0173b2",        
      "2nd Gen Combat Helicopters" = "#029e73", 
      "3rd Gen Main Battle Tanks" = "#cc78bc"   
    ),
    name = "Military Technology"
  ) +
  scale_shape_manual(
    values = c(
      "Armed UAVs" = 16,        
      "5th Gen Fighters" = 17,  
      "2nd Gen Combat Helicopters" = 15, 
      "3rd Gen Main Battle Tanks" = 18   
    ),
    name = "Military Technology"
  ) +
  scale_x_continuous(
    breaks = seq(1990, 2025, by = 5),
    limits = c(1990, 2023),
    expand = expansion(mult = c(0.01, 0.02))
  ) +
  scale_y_continuous(
    breaks = seq(0, 90, by = 10),
    expand = expansion(mult = c(0, 0.05))
  ) +
  labs(
    title = "Armed UAVs: The Fastest Military Technology Diffusion in Modern History",
    x = "Year of First Adoption",
    y = "Cumulative Number of States",
     caption = "Source: COW Arms Technology Dataset v1.0 (2025). Shaded areas show adoption trajectories. Colorblind friendly."
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5, margin = margin(b = 15)),
    legend.position = "bottom",
    legend.text = element_text(size = 11),
    legend.title = element_text(size = 12, face = "bold"),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "grey90", linewidth = 0.3),
    plot.caption = element_text(size = 8, color = "grey50", hjust = 0, margin = margin(t = 10)),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA)
  )
```




